Sketch Recognition using Particle Swarm Optimization Algorithm

### 2009-04-21 trunk/draft/Results.tex
after the grammer mistake check

\section{Datasets}
\label{sec:Datasets}
Two different data set were used, 
%For classifier and algorithm testing the system used a data set collected by
\begin{description}
	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem. The data are drawn by 16 users each of them have drawn from 50 to 60 samples for each shape. Figure \ref{fig:symbolSet} shows the shapes used in the data set. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset. In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively. %, 94.4\%  , which
%is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
		\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users. The dataset contains symbols from two different domains Electrical Engineering symbols (EL-DB) and logic design symbols (LD-DB). Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes. Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset. Symbols from different domains were gathered to evaluate the system throughout different types of shapes. Users were requested to draw symbols without any restriction on style, size or orientations. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset. %In worth noting that \cite{Oltmans07} gathered simillar dataset and acheived 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols. 
\end{description}

Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments.

@@

%We performed several experiments to evaluate the presented recognition system. Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3} in section \ref{sec:BenchMarckAlgorithm} )  to use as reference to compare it with our swarm algorithms. Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD}  shows the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. The ellipse fitting module appears to be superior. The result shows that combining both (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm. %show that combining algoirthmsachieve higher accuracy than any single algoirthm. 
%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms. 
\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{paperImages/testAlg.jpg.eps}
	 	\caption{Algorithm comparison on HS-DB} The recognition rate of different algorithms. 
	 	\label{fig:test1}
	%\caption{Experiments results :} a)    b) c) 
\end{figure} 
\begin{figure}
	\centering		
	 \includegraphics[scale=0.8]{results/AlgLD.eps}
	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
	 	\label{fig:testLD}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

@@

 The effect of symbol complexity in a dataset were testing in the following experiment. These experiments test how many samples will be required to finally obtain good performance from the recognizer system. In the following experiments, we tested the system using different number of sample per category. 
 Figure \ref{fig:Alg1Ds} and \ref{fig:Alg2Ds} shows that Hs-DB requires more number of sample for training to achieve same result as in other data sets. This is due that it contains less signal and conceptual variation that the other two datasets. Figure \ref{fig:samplesofHS-DB.jpg} shows sample of data collected from one user in Hs-Db it is clear that it has reasonable amount of variation between users but it is limited. Recognizing symbols that contains high variation in drawing style using fewer symbols in training prove the efficiency of our system.
 \begin{figure}
	\centering
		\includegraphics{results/Alg1Ds.eps}
	\caption{AlgS1 on different datasets}
	\label{fig:Alg1Ds}
\end{figure}
\begin{figure}
	\centering
		\includegraphics{results/Alg2Ds.eps}
	\caption{AlgS2 on different Datasets }
	\label{fig:Alg2Ds}
\end{figure}

  Figure \ref{} and \ref{} shows examples of correctly recognized symbol using our system. The figure shows that system recognized some hard to recognize example in both EL-DB and DL-DB. An example of symbols that was not correctly recognized are in Figure \ref{}. Most of the wrong classification were due to how messy the user draw symbols and how similarly they are conceptual. Table \ref{} show the confusion matrix of EL-DB which shows that most error were between capacitor, batter and crystal which differ only on number of parallel lines. Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallelogram and trapezoid, moon and arch. Those error were due to users drawing styles as some user may draw a parallelogram as a trapezoid.

@@

Another experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}). It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols. The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines. This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation. Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols. The penalty for this improved performance is the computational time required to run both swarm algorithms. Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols. The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles. This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them. This lead us to the next experiment to choose best set of feature to recognize symbols.

@@

\subsubsection{Features Comparisons}
\label{sec:featuresComparisions}
Different feature sets are tested to determine the best features that can used in sketch and symbol recognition. Figure\ref{fig:testFeaturesAllHS} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}). Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features. This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}. Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.  Figure \ref{fig:testFeaturesAllMD} shows the similar results on \textsl{EL-DB} and \textsl{LD-DB} dataset. The figures show that on both datasets \textbf{FS1} perform better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{EL-DB} and \textsl{LD-DB} has more structural and conceptual differences than shapes in \textsl{HS-DB}. In \cite{HeloiseBeautification} used Zernike moment(FS3) and mentioned that order higher than 8 yield to little or no improvement. In our system we used much higher order to reach the this saturation level. This is due to the level of details in the EL-DB and DL-DB sets. % achieved  HS-DB a
 \begin{figure}
	\centering
		\includegraphics[scale=0.5]{paperImages/featuresdifference.jpg.eps}
	\caption{Feature Comparison on Hs-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesAll}
\end{figure}

@@

The result illustrated in this chapter shows the work done on the system and experiments performed. We showed result of three different dataset. The system performed better than most known algorithm on the EL-DB and LD-DB. This is due to the focus on structure and stroke based properties gained from the system after correctly segmenting the input strokes. The result also showed that even using \textsl{AlgS2} only as the segmentation algorithm the system perform better than other segmentation algorithm. Using different databases shows that the more the structure difference in the symbols the better recognition rate. LD-DB proved to have the lowest recognition rate due to the high similarity between its symbols.    % better than

>>>>

\section{Datasets}
\label{sec:Datasets}
Three different data set were used, table \ref{tab:datasets} shows a comparison between the three datasets used in our experiments.

%For classifier and algorithm testing the system used a data set collected by
\begin{description}
	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem. The data are drawn by 16 users each of them have drawn from 50 to 60 samples for each shape. Figure \ref{fig:symbolSet} shows the shapes used in the data set. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset. In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively. %, 94.4\%  , which
%is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
		\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users. The dataset contains symbols from two different domains Electrical Engineering symbols (EL-DB) and logic design symbols (LD-DB). Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes. Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset. Symbols from different domains were gathered to evaluate the system throughout different types of shapes. Users were requested to draw symbols without any restriction on style, size or orientations. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset. %In worth noting that \cite{Oltmans07} gathered similar dataset and achieved 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols. 
\end{description}

@@

%We performed several experiments to evaluate the presented recognition system. Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3} in section \ref{sec:BenchMarckAlgorithm} )  to use as reference to compare it with our swarm algorithms. Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD}  shows the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. The ellipse fitting module appears to be superior. The result shows that combining both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) outperform any single algorithm. %show that combining algorithms achieve higher accuracy than any single algoirthm. 
%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms. 
\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{paperImages/testAlg.jpg.eps}
	 	\caption{Algorithm comparison on HS-DB} The recognition rate of different algorithms. 
	 	\label{fig:test1}
	%\caption{Experiments results :} a)    b) c) 
\end{figure} 
\begin{figure}
	\centering		
	 \includegraphics[scale=0.8]{results/AlgLD.eps}
	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
	 	\label{fig:testLD}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

@@

 The effects of symbol complexity in a dataset were testing in the following experiment. These experiments test how many samples will be required to finally obtain good performance from the recognizer system. In the following experiments, we tested the system using different number of sample per category. 
 Figure \ref{fig:Alg1Ds} and \ref{fig:Alg2Ds} shows that Hs-DB requires more number of sample for training to achieve same result as in other data sets. This is due that it contains less signal and conceptual variation that the other two datasets. Figure \ref{fig:samplesofHS-DB.jpg} shows sample of data collected from one user in Hs-Db it is clear that it has reasonable amount of variation between users but it is limited. Recognizing symbols that contains high variation in drawing style using fewer symbols in training prove the efficiency of our system.
 \begin{figure}
	\centering
		\includegraphics{results/Alg1Ds.eps}
	\caption{AlgS1 on different datasets}
	\label{fig:Alg1Ds}
\end{figure}
\begin{figure}
	\centering
		\includegraphics{results/Alg2Ds.eps}
	\caption{AlgS2 on different Datasets }
	\label{fig:Alg2Ds}
\end{figure}

  Figure \ref{} and \ref{} shows examples of correctly recognized symbol using our system. The figure shows that system recognized some hard to recognize example in both EL-DB and DL-DB. An example of symbols that was not correctly recognized are in Figure \ref{}. Most of the wrong classifications were due to how messy the user draw symbols and how similarly they are conceptual. Table \ref{} show the confusion matrix of EL-DB which shows that most error were between capacitor, batter and crystal which differ only on number of parallel lines. Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallelogram and trapezoid, moon and arch. Those errors were due to users drawing styles as some user may draw a parallelogram as a trapezoid.

@@

Another experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}). It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.  The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines. This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation. Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols. The penalty for this improved performance is the computational time required to run both swarm algorithms. Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols. The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles. This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them. This led us to the next experiment to choose best set of feature to recognize symbols.

@@

\subsubsection{Features Comparisons}
\label{sec:featuresComparisions}
Different feature sets are tested to determine the best features that can used in sketch and symbol recognition. Figure\ref{fig:testFeaturesAllHS} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}). Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features. This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}. Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.  Figure \ref{fig:testFeaturesAllMD} shows the similar results on \textsl{EL-DB} and \textsl{LD-DB} dataset. The figures show that on both datasets \textbf{FS1} perform better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{EL-DB} and \textsl{LD-DB} has more structural and conceptual differences than shapes in \textsl{HS-DB}. In \cite{HeloiseBeautification} used Zernike moment (FS3) and mentioned that order higher than 8 yield to little or no improvement. In our system we used much higher order to reach this saturation level. This is due to the level of details in the EL-DB and DL-DB sets. % achieved  HS-DB a
 \begin{figure}
	\centering
		\includegraphics[scale=0.5]{paperImages/featuresdifference.jpg.eps}
	\caption{Feature Comparison on Hs-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesAll}
\end{figure}

@@

The result illustrated in this chapter shows the work done on the system and experiments performed. We showed result of three different dataset. The system performed better than most known algorithm on the EL-DB and LD-DB. This is due to the focus on structure and stroke based properties gained from the system after correctly segmenting the input strokes. The result also showed that even using \textsl{AlgS2} only as the segmentation algorithm the system perform better than other segmentation algorithm. Using different databases shows that the more the structure differences in the symbols the better recognition rate. LD-DB proved to have the lowest recognition rate due to the high similarity between its symbols.    % better than
